---
title: "Data Description for FRST 556 course data"
editor: visual
---

```{r, echo = FALSE}
yml_content <- yaml::read_yaml("chapterauthors.yml")
author <- yml_content[["terrain-spatial-interpolation"]][["author"]]
```

```{r, include = FALSE, echo = FALSE}
# Check if leaflet and tidyverse are installed using pacman package
if (!require(pacman)) install.packages("pacman")
pacman::p_load(leaflet, tidyverse)
```

## Course Data {#sec-data}

Written by

```{r results='asis', echo=FALSE}
cat(author)
```

FRST 556 is mostly concentrated on Malcom Knapp Research Forest (MKRF) located outside of Maple Ridge, British Columbia. The research forest maintains an extensive GIS database available to researchers and students alike. We have included the necessary files in the material posted on the UBC canvas page. Casual users can access the data here: [Access MKRF Data](https://www.mkrf.forestry.ubc.ca/education/gis/)

```{r, include = FALSE}
# Coordinates for Malcolm Knapp Research Forest
lat <- 49.2833
lng <- -122.5833

# Create a leaflet map with a satellite basemap
leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  setView(lng = lng, lat = lat, zoom = 14) %>%
  addMarkers(lng = lng, lat = lat, popup = "Malcolm Knapp Research Forest")
```

## Data Incorporated in Course Material

The original GIS material available from MKRF have been converted from the ESRI geodatabase format into a series of geopackages. In the folder, the first letter describes the datatype. These are described in the table below in addition to the data included

+--------------+---------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Start Letter | Type    | Description                                                                                                                                                      |
+==============+=========+==================================================================================================================================================================+
| I            | Image   | I_2006_MKRF_RGB.ecw - 2006 aerial RGB image                                                                                                                      |
|              |         |                                                                                                                                                                  |
|              |         | I_2008_MKRF_RGB.tif - 2008 mosaiced aerial RGB image                                                                                                             |
|              |         |                                                                                                                                                                  |
|              |         | ![]()I_2019_LandsatBAP_MKRF-01.tif - 2019 '[Best Available Pixel](https://code.earthengine.google.com/e27240a92ecf64bbadf8a082b91c711c?hideCode=true)' composite |
|              |         |                                                                                                                                                                  |
|              |         | I_Lake_Mask_MKRF.tif - Binary mask of Lakes within MKRF Boundaer                                                                                                 |
+--------------+---------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| L            | Line    | L_contours_1m.gpkg - 1 m elevation contours                                                                                                                      |
|              |         |                                                                                                                                                                  |
|              |         | L_contours_5m.gpkg - 5 m elevation contours                                                                                                                      |
|              |         |                                                                                                                                                                  |
|              |         | L_hydroline.gkpg - Location of hydroline crossing MKRF                                                                                                           |
|              |         |                                                                                                                                                                  |
|              |         | L_roads_y2010.gpkg - Location of roads (as of 2010)                                                                                                              |
|              |         |                                                                                                                                                                  |
|              |         | L_streams_major.gkpg - Major streams                                                                                                                             |
|              |         |                                                                                                                                                                  |
|              |         | L_streams.gkpg - all streams                                                                                                                                     |
|              |         |                                                                                                                                                                  |
|              |         | L_trails.gkpg - Location of trails within MKRF                                                                                                                   |
+--------------+---------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| P            | Polygon | P_BC_Biogeoclimatic.gpkg - Biogeoclimatic areas of MRKF                                                                                                          |
|              |         |                                                                                                                                                                  |
|              |         | P_BC_VRI_2014.gpkg - 2014 VRI based on 1996 aerial imagery and then updated (i.e. projected) to 2014                                                             |
|              |         |                                                                                                                                                                  |
|              |         | P_Boundary_MKRF.gpkg - MKRF boundary                                                                                                                             |
|              |         |                                                                                                                                                                  |
|              |         | P_BoundaryWoodlot.gpkg - Boundary for woodlot adjacent to MKRF also managed by MKRF staff                                                                        |
|              |         |                                                                                                                                                                  |
|              |         | P_Buildings.gpkg - Location of structures for Loon Lake Retreat Center                                                                                           |
|              |         |                                                                                                                                                                  |
|              |         | P_ForestCover_1989.gpkg - Photo- interpretation of forest cover from BC sterography                                                                              |
|              |         |                                                                                                                                                                  |
|              |         | P_ForestCover_1999.gpkg - Updated version of the 1989 assessment based on known locations of human and natural disturbances                                      |
|              |         |                                                                                                                                                                  |
|              |         | P_ForestCover_2008.gpkg - Updated version of 1989 and 1999 assessment based on known location of human and natural disturbances                                  |
|              |         |                                                                                                                                                                  |
|              |         | P_ForestDevelopmentPlan_2019.gpkg - Location of forest development plans as per 2019                                                                             |
|              |         |                                                                                                                                                                  |
|              |         | P_lakes.gpgk - Polygon of lakes in MKRF                                                                                                                          |
|              |         |                                                                                                                                                                  |
|              |         | P_proposedharvest_y2020.gpkg - Location of proposed harvests as of 2020                                                                                          |
|              |         |                                                                                                                                                                  |
|              |         | P_proposedharvest_y2023.gpkg - Location of proposed harvests as of 2023                                                                                          |
|              |         |                                                                                                                                                                  |
|              |         | P_ReserveAreas.gpkg - Location of reserves where generally no harvest occur (except for safety issues)                                                           |
+--------------+---------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Pts          | Points  | Pts_SelectiveSampling_2016.gpkg - Locations of selective sampling for 2016                                                                                       |
|              |         |                                                                                                                                                                  |
|              |         | Pts_SystematiceSampling_1995.gpkg - Locations of the systematic sampling areas                                                                                   |
+--------------+---------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+

: Description of Files included in the 'Data_Folder'

# Field Code Descriptions for "Forest Cover" Layers {#sec-fieldcodes}

| Type                 | Field                   | Description                                                                                                                                                                                                                                                                                                                 |
|----------------------|-------------------------|-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| Description          | SHAPE                   | Internally ArcView                                                                                                                                                                                                                                                                                                          |
| Description          | Fc99#                   | Internally Arc/INFO                                                                                                                                                                                                                                                                                                         |
| Description          | Fc99-id                 | Internally Arc/INFO                                                                                                                                                                                                                                                                                                         |
| Description          | Poly#                   | Internally Arc/INFO                                                                                                                                                                                                                                                                                                         |
| Description          | Subclass                | Internally Arc/INFO                                                                                                                                                                                                                                                                                                         |
| Description          | Subclass#               | Internally Arc/INFO                                                                                                                                                                                                                                                                                                         |
| Description          | Rings_ok                | Internally Arc/INFO                                                                                                                                                                                                                                                                                                         |
| Description          | Rings_nok               | Internally Arc/INFO                                                                                                                                                                                                                                                                                                         |
| Description          | ID                      | Polygon ID – unique for every polygon in the map                                                                                                                                                                                                                                                                            |
| Description          | AREA                    | Polygon area                                                                                                                                                                                                                                                                                                                |
| Description          | PERIMETER               | Polygon perimeter                                                                                                                                                                                                                                                                                                           |
| Description          | MAPNO                   | Map number (1, 2 or 3)                                                                                                                                                                                                                                                                                                      |
| Description          | UID                     | Unit ID – Combination of map sheet number and polygon ID                                                                                                                                                                                                                                                                    |
| Description          | UID2                    | Unit ID Specification – range from “a” to “z” and is due to the process of splitting the initial polygons by the partial cuttings                                                                                                                                                                                           |
| Description          | NAME                    | Ex. 97A80_01 97-Date of treatment, A80-name of road, \_01- \# of cutblock along road                                                                                                                                                                                                                                        |
| Description          | TREATMENT               | Silvicultural Treatment (Clearcut or Thinning)                                                                                                                                                                                                                                                                              |
| Description          | LAYER                   | Horizontal stratum or layer: 1 - most important, 2 - next most important layer                                                                                                                                                                                                                                              |
| Description          | RANK                    | Order of importance of each layer in a multi-layered stand                                                                                                                                                                                                                                                                  |
| Description          | NFD                     | Non-forest descriptor- **C** – cultivated- **LK** – lake- **NCBR** – non-commercial brush- **NPBR** – non-productive brush- **NP** – non-productive- **NSR** – non-sufficiently restocked- **RIV** – river- **SW** –- **TS** – mine tailing prevent growth- **U** - urban                                                   |
| Description          | DCLASS                  | Data collection class: 0 - photo interpretation, 1- air call, 8 - ground observation                                                                                                                                                                                                                                        |
| Description          | DSOURCE                 | Data collection source: 0 - no qualification, 1 - Ionut Aron, 2 - MRKF staff                                                                                                                                                                                                                                                |
| Description          | DATE_IN                 | Data of the initial forest inventory (YYMMDD)                                                                                                                                                                                                                                                                               |
| Description          | DATE_PROJ               | Date for any projected fields (YYMMDD)                                                                                                                                                                                                                                                                                      |
| Photo interpretation | SPCS1… SPCS6            | Species from the most prevalent (**SPSC1**)- **AC** – Balsam poplar- **H** – Western hemlock- **F** – Douglas-fir- **BG** – Grand fir- **D** – Red alder- **C** – Western redcedar- **L** – Larch- **MB** – Broadleaf maple- **CY** – Yellow cedar- **PL** – Lodgepole pine- **PW** – Western white pine                    |
| Photo interpretation | PCT1 … PCT6             | Percentages for **SPCS1** – **SPCS6**                                                                                                                                                                                                                                                                                       |
| Photo interpretation | AGECL_IN                | Initial age class at the time of the forest inventory- **1** - \<20 years- **2** – 21-40 years- **3** – 41-60 years- **4** – 61-80 years- **5** – 81-100 years- **6** – 101-120 years- **7** – 121-140 years- **8** – 141-250 years- **9** - \> 250 years                                                                   |
| Photo interpretation | HTCL_IN                 | Initial height class at the time of the forest inventory- **1** - \< 10.4 m- **2** – 10.5-19.4 m- **3** – 19.5-28.4 m- **4** – 28.5-37.4 m- **5** – 37.5-46.4 m- **6** – 46.5-55.4 m- **7** – 55.5-64.4 m- **8** - \>64.5 m                                                                                                 |
| Photo interpretation | SITE_IN                 | Site index class- **L** – low- **P** – poor- **M** – medium- **G** – good                                                                                                                                                                                                                                                   |
| Photo interpretation | CRNCLCL                 | Crown closure class- **0** – 0-5%- **1** – 6-15%- **2** – 16-25%- **3** – 26-35%- **4** – 36-45%- **5** – 46-55%- **6** – 56-65%- **7** – 66-75%- **8** – 76-85%- **9** – 86-95%- **10** – 96-100%                                                                                                                          |
| Photo interpretation | BASICL                  | Basic Class – the class or type of non-productive areas- **0** – Forest Cover- **3** – Rock- **4** – This code does not appear in the Ministry of Forests’ documentation- **11** – Non-productive brush- **12** – Non-productive- **15** – Lake- **25** – River- **35** – Wwamp (muskeg)- **42** – Clearing- **54** – Urban |
| Photo interpretation | EPA                     | Environmentally sensitive areas- **ER** – High recreation- **RW** – Recreation and wildlife- **W** – wildlife- **S** – unstable                                                                                                                                                                                             |
| Derived Values       | AGE_PLANTE              | Age of trees when planted                                                                                                                                                                                                                                                                                                   |
| Derived Values       | HEIGHT_IN               | Initial average height (m)                                                                                                                                                                                                                                                                                                  |
| Derived Values       | AGE_IN                  | Initial stand age                                                                                                                                                                                                                                                                                                           |
| Derived Values       | SITEINDX                | VDYP site index                                                                                                                                                                                                                                                                                                             |
| Derived Values       | VOL12_5                 | Average stand volume above 12.5 cm diameter limit                                                                                                                                                                                                                                                                           |
| Derived Values       | VOL7_5                  | Average stand volume above 7.5 cm diameter limit                                                                                                                                                                                                                                                                            |
| Derived Values       | DIAM12_5                | Average stand diameter above 12.5 cm diameter limit                                                                                                                                                                                                                                                                         |
| Derived Values       | DIAM7_5                 | Average stand diameter above 7.5 cm diameter limit                                                                                                                                                                                                                                                                          |
| Derived Values       | CRNCLOS                 | Crown closure (%)                                                                                                                                                                                                                                                                                                           |
| Derived Values       | DENSITY                 | Density (stems per hectare                                                                                                                                                                                                                                                                                                  |
| Derived Values       | REFYR                   | Reference year for update                                                                                                                                                                                                                                                                                                   |
| Derived Values       | AGE_PROJ                | Projected age – range from 0 to 500 years                                                                                                                                                                                                                                                                                   |
| Derived Values       | HEIGHT_PRO              | Projected height (m) – range from 0 to 50 m                                                                                                                                                                                                                                                                                 |
| Derived Values       | AGECL_PROJ              | Projected age class (same as Initial age class coding) – range from 1 to 9                                                                                                                                                                                                                                                  |
| Derived Values       | HTCL_PROJ               | Projected height class (same as Initial height class coding) – range from 1 to 8                                                                                                                                                                                                                                            |
| Derived Values       | DATE_EST                | Date forest established calculated by subtracting the mean age of the leading species of the stand or plot from the calendar year that the measurement or estimate was made                                                                                                                                                 |
| Ground Data          | DBH_LIM                 | Diameter at Breast Height Limit – minimum diameter at breast height to which trees are to be recorded or measured (all records have value of 22.5 cm)                                                                                                                                                                       |
| Ground Data          | BAGE                    | Age at breast height – the number of annual rings counted at 1.30 m height                                                                                                                                                                                                                                                  |
| Ground Data          | BAGECORR                | Breast age corrected to total age                                                                                                                                                                                                                                                                                           |
| Ground Data          | TOPHT                   | Top Height – average height of 100 trees of largest diameter per hectare                                                                                                                                                                                                                                                    |
| Ground Data          | GSPCS1 … GSPCS3         | Same as **SPCS1** … **SPCS6** but data from ground observations                                                                                                                                                                                                                                                             |
| Ground Data          | GPCT1 … GPCT3           | Same as **PCT1** … **PCT6** but data from ground observations                                                                                                                                                                                                                                                               |
| Ground Data          | BA_HA_SP1 …BA_HA_SP3    | Basal area per hectare for **SPCS1** … **SPCS6**                                                                                                                                                                                                                                                                            |
| Ground Data          | SITE_H_SP1 … SITE_H_SP3 | The height for the site tree per species                                                                                                                                                                                                                                                                                    |
